\chapter*{2. Probability}

\section*{2.2. Rules for three events}
\begin{enumerate}[(a)]
    \item Consider three events, $A$, $B$, and $C$. 
        If they are all independent of each other,
        show that
        \begin{equation*}\tag{2.76}
            P(A \text{ and } B \text{ and } C) = P(A) \cdot P(B) \cdot P(C) 
        \end{equation*}

    \item If they are (possibly) dependent, show that 
        \begin{equation*}\tag{2.77}
            P(A \text{ and } B \text{ and } C) = P(A) \cdot P(B | A) \cdot P(C | A \text{ and } B) 
        \end{equation*}

    \item If they are all mutually exclusive, show that
        \begin{equation*}\tag{2.78}
            P(A \text{ or } B \text{ or } C) = P(A) + P(B) + P(C)
        \end{equation*}

    \item If they are (possibly) nonexclusive, show that
        \begin{align*}\tag{2.79}
            P(A \text{ or } B \text{ or } C) =& P(A) + P(B) + P(C) \\
                                          &- P(A \text{ and } B) - P(A \text{ and } C) - P(B \text{ and } C) \\
                                          &+ P(A \text{ and } B \text{ and } C)
        .\end{align*}
\end{enumerate}

\begin{proof}
    \hfill
    \begin{enumerate}[(a)]
        \item Using (2.9) we find that:
            \[
                P(A \text{ and } B \text{ and } C) = P(A) \cdot P(B \text{ and } C | A) 
                = P(A) \cdot P(B | A) \cdot P(C | A \text{ and } B)
            .\] 
            Because the events are independent, $P(B | A) = P(B)$ and $P(C | A \text { and } B) = P(C)$, so
            \begin{equation*}\tag{2.76}
                P(A \text{ and } B \text{ and } C) = P(A) \cdot P(B) \cdot P(C). 
            \end{equation*}

        \item Proved at (a).

        \item By (2.18), 
            \begin{align*}
                P(A \text{ and } B \text{ and } C) 
                    &= P(A) + P(B \text{ or } C) - P(A \text{ and } (B \text{ or } C)) \\
                    &= P(A) + P(B) + P(C) - P(B \text{ and } C) - P(A) \cdot P(B \text{ or } C | A)
            .\end{align*}

            We compute the subtracted member:
            \begin{align*}
                P(A) \cdot P(B \text{ or } C | A) 
                    &= P(A) \cdot (P(B|A) + P(C|A) - P(B \text{ and } C | A))) \\
                    &= P(A \text{ and } B) + P(A \text{ and } C) - P(A \text{ and } B \text{ and } C)
            .\end{align*}

            By substituting in the initial expression, we get:
            \begin{align*}\tag{2.79}
                P(A \text{ or } B \text{ or } C) =& P(A) + P(B) + P(C) \\
                                              &- P(A \text{ and } B) - P(A \text{ and } C) - 
                                                P(B \text{ and } C) \\
                                              &+ P(A \text{ and } B \text{ and } C)
            .\end{align*}

            Since the events are mutually exclusive, all $\emph{and}$ probabilities are 0, so:
            \begin{equation*}\tag{2.78}
                P(A \text{ or } B \text{ or } C) = P(A) + P(B) + P(C)
            \end{equation*}

        \item Proved at (c).
    \end{enumerate}
\end{proof}

\section*{2.7. Proofreading}
Two people each proofread the same book. One person finds 100 errors, and the other finds 60. 
There are 20 errors common to both people. Assume that all errors are equally likely to be found (which is 
undoubtedly not true in practice), and also that the discovery of an error by a person is independent
of the discovery of that error by the other person. Given these assumptions, roughly how many error 
does the book have? $\emph{Hint:}$ Draw the picture similar to Fig. 2.1, and then find the probability 
of each person finding a given error.

\vspace{2em}

\begin{proof}
    \hfill

    Let $P(A)$ be the probability that a random error is one of the errors discovered 
    by the person with 100 errors, and respectively let $P(B)$ be the same for the person with 60 errors.
    Also, let $N$ be the approximate number of total errors.
    Since the discoveries of errors are independent events, 
    \[
        P(A \text{ and } B) = P(A) \cdot P(B) = \frac{100}{N} \cdot \frac{60}{N} = \frac{6000}{N^2}
    .\] 
    But we know that 20 errors are common between the two persons, so
    \[
        P(A \text{ and } B) = \frac{20}{N}
    .\] 
    Therefore, we get the approximated total number of errors:
    \[
        \frac{20}{N} = \frac{6000}{N^2} \iff N = 300
    .\] 
\end{proof}
